---
title: "activity2.Rmd"
output: github_document
---

```{r echo=FALSE, cache= F}
#Source.R file has package installation and tableFormat function 
source('source.R')
```

###Join, merge, look up

####Activity #2

Creating datasets (movies and studios) for explaining the joins
```{r}
names<-c("Iron Man","Thor","The Avengers","Ant Man","Doctor Strange","Man of steel","Dawn of Justice","Suicide squad","Wonder woman","Justice league","Force awakens")
release_year<-c("2008","2008","2011","2015","2016","2013","2016","2016","2017","2017","2015")
directors<-c("Jon Favreau","Kenneth Branagh","Joss Whedon","Peyton Reed","Scott Derrickson","Zack Snyder","Zack Snyder","David Ayer","Patty Jenkins","Zack Snyder","J.J. Abrams")
company<-c(rep("Marvel",5),rep("DC",5),"Star wars")
movies<-data.frame(names,release_year,directors,company)

company<-c("Marvel","DC","Star trek")
yr_founded<-c(1934,1939,1966)
studios<-data.frame(company,yr_founded)
```

Lets view the datasets movies 

```{r}
tableFormat(movies)
```

Lets view the datasets studios
```{r}
tableFormat(studios)
```

Lets explore the different ways of joining these two dataframes.

###Different types of join

####inner_join(movies,studios)

> inner_join(x, y): Return all rows from x where there are matching values in y, and all columns from x and y. If there are multiple matches between x and y, all combination of the matches are returned. This is a mutating join.

```{r}
#inner_join
suppressWarnings(suppressMessages(
  tableFormat(inner_join(movies,studios))
))
```
We lose Force Awakens in the join because, although they appear in x = movies, their corrosponding company "Star wars" doesn't not appear in y = studios. The join result has all variables from x = movies plus yr_founded, from y.

####semi_join(movies,studios)

> semi_join(x, y): Return all rows from x where there are matching values in y, keeping just columns from x. A semi join differs from an inner join because an inner join will return one row of x for each matching row of y, where a semi join will never duplicate rows of x. This is a filtering join.

```{r}
#semi join
suppressWarnings(suppressMessages(
  tableFormat(semi_join(movies,studios))
))
```
We get a similar result as with inner_join() but the join result contains only the variables originally found in  x = movies. But note the row order has changed.

####left_join(movies,studios)

> left_join(x, y): Return all rows from x, and all columns from x and y. If there are multiple matches between x and y, all combination of the matches are returned. This is a mutating join.

```{r}
#left join
suppressWarnings(suppressMessages(
  tableFormat(left_join(movies,studios))
))
```

We basically get x = movies back, but with the addition of variable yr_founded, which is unique to  y = studios. Force Awakens, whose company don't appear in y = studios, has an NA for company.

####anti_join(movies,studios)

> anti_join(x, y): Return all rows from x where there are not matching values in y, keeping just columns from x. This is a filtering join.

```{r}
#anti join
suppressWarnings(suppressMessages(
  tableFormat(anti_join(movies,studios))
))
```
We keep only Force Awakens now (and do not get company).

####inner_join(studios,movies)

> inner_join(x, y): Return all rows from x where there are matching values in y, and all columns from x and y. If there are multiple matches between x and y, all combination of the matches are returned. This is a mutating join.

```{r}
#inner_join
suppressWarnings(suppressMessages(
  tableFormat(inner_join(studios,movies))
))
```
In a way, this does illustrate multiple matches, if you think about it from the x = studios direction. Every company that has a match in y = movies appears multiple times in the result, once for each match. In fact, we’re getting the same result as with inner_join(movies,studios), up to variable order (which you should also never rely on in an analysis).

####semi_join(studios,movies)

> semi_join(x, y): Return all rows from x where there are matching values in y, keeping just columns from x. A semi join differs from an inner join because an inner join will return one row of x for each matching row of y, where a semi join will never duplicate rows of x. This is a filtering join.

```{r}
#semi join
suppressWarnings(suppressMessages(
  tableFormat(semi_join(studios,movies))
))
```
Now the effects of switching the x and y roles is more clear. The result resembles x = studios, but the company star wars is lost, because there are no observations where company == "star wars" in y = movies.

####left_join(studios,movies)

> left_join(x, y): Return all rows from x, and all columns from x and y. If there are multiple matches between x and y, all combination of the matches are returned. This is a mutating join.

```{r}
#left join
suppressWarnings(suppressMessages(
  left_join(studios,movies)
))
```

We get a similar result as with inner_join() but the company star wars survives in the join, even though no moveis from star wars that appear in y = movies. As a result, star wars has NA in names,release_year and directors.

####anti_join(studios,moviess)

> anti_join(x, y): Return all rows from x where there are not matching values in y, keeping just columns from x. This is a filtering join.

```{r}
#anti join
suppressWarnings(suppressMessages(
  tableFormat(anti_join(studios,movies))
))
```
We keep only company star wars now (and the variables found in x = studios).

####full_join

> full_join(x, y): Return all rows and all columns from both x and y. Where there are not matching values, returns NA for the one missing. This is a mutating join.

```{r}
#full join
suppressWarnings(suppressMessages(
  tableFormat(full_join(movies,studios))
))
```

We get all rows of x = movies plus a new row from y = movies, containing the company star wars. We get all variables from x = movies AND all variables from y = studios. Any row that derives solely from one table or the other carries NAs in the variables found only in the other table.
