---
title: "hw01_gapminder.Rmd"
output: github_document
---

First install and load packages gapminder and tidyverse
```{r}
#install.packages("gapminder")
#install.packages("tidyverse")
library(gapminder)
library(tidyverse)
```

Check the nature of gapminder data using str()

```{r}
str(gapminder)
```
Check first few rows using head() and and last few rows using tail()
```{r}
head(gapminder)
tail(gapminder)
```

Run basic queries on gapminder data frame
```{r}
names(gapminder)
ncol(gapminder)
length(gapminder)
dim(gapminder)
nrow(gapminder)
```

Get an overview of gapminder data using summary() function
```{r}
summary(gapminder)
```

#####Few basic plots

Plot of Life Experience against Year
```{r}
plot(lifeExp ~ year, gapminder)
```

Plot of Life Experience against Gdp per capta
```{r}
plot(lifeExp ~ gdpPercap, gapminder)
```

######Lets do little transformation on Gdp Per cap

Here's a plot of Life Experience against log of Gdp per capta
```{r}
plot(lifeExp ~ log(gdpPercap), gapminder)
```

######Now, lets look at specific variables inside the gapminder data frame.

First lets expore the numeric variable for life Expectancy (lifeExp)
```{r}
head(gapminder$lifeExp)
summary(gapminder$lifeExp)
hist(gapminder$lifeExp)
```

Secondly, lets know more about the Integer variable year
```{r}
summary(gapminder$year)
table(gapminder$year)
```

######Categorical variables are stored as factor in R. 
Now, Lets expore the continent factor variable from gapminder data.
```{r}
class(gapminder$continent)
summary(gapminder$continent)
levels(gapminder$continent)
nlevels(gapminder$continent)
```

*But R is really storing integer codes for the different levels of factor variables.*
This can be seen from result of str() function.
```{r}
str(gapminder$continent)
```

Visual portrayal of continent data using table and barplot function
```{r}
table(gapminder$continent)
barplot(table(gapminder$continent))
```

