---
title: "activity1.Rmd"
output: github_document
---

Install all the packages

```{r}
suppressPackageStartupMessages(library(tidyverse))
suppressPackageStartupMessages(library(gapminder))
suppressPackageStartupMessages(library(ggplot2))
suppressPackageStartupMessages(library(knitr))
suppressPackageStartupMessages(library(kableExtra))
suppressPackageStartupMessages(library(tidyr))
suppressPackageStartupMessages(library(dplyr))
suppressPackageStartupMessages(library(reshape2))
suppressPackageStartupMessages(library(gridExtra))
options(knitr.table.format = "html")
```

User defined Function for formatting tables
```{r}
tableFormat<-function(table,title=""){
  table %>%      
    kable("html",caption=title, align=c(rep('c', 5))) %>%
    kable_styling(bootstrap_options = 
                    c("striped", "hover",  "responsive"),
                  position="center",font_size=14)
}

```

###General data reshaping and relationship to aggregation

####Activity #1 

Lets explore the important tidyr functions and how they are similar to the dplyr functions.

Lets build a data for our purpose.. 
```{r}
dat <- data.frame(
  id = LETTERS[1:5],
   x = sample(0:1, 5, TRUE),
   y = sample(0:1, 5, TRUE),
   z = sample(0:1, 5, TRUE),
   category = c("N1","N2","N3","N4","N5"),
   score = c(134.54,123.23,112.38,129.5,110.8)
)
tableFormat(dat,"Sample Data")
```


gather()

Lets explore the function gather().

```{r}
a<- dat%>%
      gather(coordinate,values,-c(category,id)) #dont gether category and id
tableFormat(a)
```

```{r}
b<- dat%>%
      gather(coordinate,values,-c(category,id,score))
tableFormat(b)
```

```{r}
c<- dat%>%
      gather(coordinate,values,x:z,-y)   #gathering only x&z
tableFormat(c)
```

**spread()**
This does the opposite of the function gather().

```{r}
d<- dat%>%
      gather(coordinate,values,x,y)%>%
      spread(coordinate,values)
tableFormat(d)
```

```{r}
e<- dat%>%
      gather(coordinate,values,-c(category,id,score))
tableFormat(e)
```

```{r}
f<- dat%>%
      gather(coordinate,values,-c(category,id,score))%>%
      spread(coordinate,values)%>%
      spread(category,score,fill=0)
tableFormat(f)
```

Note the re ordering of columns after performing gathering and spreading operations simultaneously.

Now, lets look at **extract()**,**complete()**, **expand()** and **replace_na()**

```{r}
#taking no. in category & get variable class
g<- dat%>%
      extract(category,"class",regex="([0-9]+)")
tableFormat(g)  
```

```{r}
#Creates rows for each id for each score in the data.
h<- dat%>%
      complete(category, nesting(id),score)
tableFormat(h)  
```

```{r}
# complete is for a common application of expand
i<- dat%>%
      expand(category,nesting(id),score)
tableFormat(i) 
```

complete() 
> completing a data frame with missing combinations.

expand()
> Expanding a data frame to include all combinations.

```{r}

j<- dat%>%
      complete(category, nesting(id,score))%>% #Finds combinations of id & score
      replace_na(list(x=-1,y=-1,z=-1))        #Replaces nas with -1 in x,y,z.
tableFormat(j) 
```

**separate()** and **unite()** functions
Lets explore the functions separate() and unite().

```{r}
l<- dat%>%
      unite(idAndCategory,id,category,sep=":",remove = FALSE)%>%
      select(-id,-category)%>%
      separate(idAndCategory, c("id", "category"), ":")
tableFormat(l)  
```

**unnest()** function
Now lets check the unnest() function.

```{r}
#makes each element of list column its own row.
m<- dat%>%
      unite(idAndCategory,id,category,sep=":",remove = FALSE)%>%
      select(-id,-category)%>%
      unnest(idAndCategory=strsplit(idAndCategory, ":"))
tableFormat(m)   
```

