---
title: "activity1.Rmd"
output: github_document
---

Install all the packages

```{r}
suppressPackageStartupMessages(library(tidyverse))
suppressPackageStartupMessages(library(gapminder))
suppressPackageStartupMessages(library(ggplot2))
suppressPackageStartupMessages(library(knitr))
suppressPackageStartupMessages(library(kableExtra))
suppressPackageStartupMessages(library(tidyr))
suppressPackageStartupMessages(library(dplyr))
suppressPackageStartupMessages(library(reshape2))
suppressPackageStartupMessages(library(gridExtra))
options(knitr.table.format = "html")
```

User defined Function for formatting tables
```{r}
tableFormat<-function(table,title=""){
  table %>%      
    kable("html",caption=title, align=c(rep('c', 5))) %>%
    kable_styling(bootstrap_options = 
                    c("striped", "hover",  "responsive"),
                  position="center",font_size=14)
}

```

###General data reshaping and relationship to aggregation

####Activity #1 

Lets explore the important tidyr functions and how they are similar to the dplyr functions.

Lets build a data for our purpose.. 
```{r}
dat <- data.frame(
  id = LETTERS[1:5],
   x = sample(0:1, 5, TRUE),
   y = sample(0:1, 5, TRUE),
   z = sample(0:1, 5, TRUE),
   category = c("N1","N2","N3","N4","N5"),
   score = c(134.54,123.23,112.38,129.5,110.8)
)
tableFormat(dat,"Sample Data")
```


**gather()**  

Lets explore the function gather().

```{r}
dat%>%
  gather(coordinate,values,-c(category,id))%>% #dont gether category and id
  tableFormat()
  
dat%>%
  gather(coordinate,values,-c(category,id,score))%>%
  tableFormat()

dat%>%
  gather(coordinate,values,x:z,-y)%>%   #gathering only x&z
  tableFormat()

#gather(),separate(),spread(),extract()
```


**spread()**
This does the opposite of the function gather().

```{r}
dat%>%
  gather(coordinate,values,x,y)%>%
  spread(coordinate,values)%>%
  tableFormat()

dat%>%
  gather(coordinate,values,-c(category,id,score))%>%
  tableFormat()

dat%>%
  gather(coordinate,values,-c(category,id,score))%>%
  spread(coordinate,values)%>%
  spread(category,score,fill=0)%>%
  tableFormat()
```

Note the re ordering of columns after performing gathering and spreading operations simultaneously.

Now, lets look at **extract()**,**complete()**, **expand()** and **replace_na()**

```{r}
#taking no. in category & get variable class
dat%>%
  extract(category,"class",regex="([0-9]+)")%>%
  tableFormat()  

#Creates rows for each id for each score in the data.
dat%>%
  complete(category, nesting(id),score)%>%
  tableFormat()  
  
# complete is for a common application of expand
dat%>%
  expand(category,nesting(id),score)%>%
  tableFormat()  

#complete() - completing a data frame with missing combinations.
#expand() - Expanding a data frame to include all combinations.

dat%>%
  complete(category, nesting(id,score))%>%
  #Finds the combinations of id and score that occur in the data.
  replace_na(list(x=-1,y=-1,z=-1))%>%
  #Replaces the na values with -1 in x,y,z.
  tableFormat()  
```

**separate()** and **unite()** functions
Lets explore the functions separate() and unite().

```{r}
dat%>%
  unite(idAndCategory,id,category,sep=":",remove = FALSE)%>%
  select(-id,-category)%>%
  separate(idAndCategory, c("id", "category"), ":")%>%
  tableFormat()  
```

**unnest()** function
Now lets check the unnest() function.

```{r}
#makes each element of list column its own row.
dat%>%
  unite(idAndCategory,id,category,sep=":",remove = FALSE)%>%
  select(-id,-category)%>%
  unnest(idAndCategory=strsplit(idAndCategory, ":"))%>%
  tableFormat()   

```

